A Horrocks-type theorem for even orthogonal $K_2$

### 2020-06-13 dl-paper.tex
Fix misprint
Fix misprints

\begin{lemma} \label{lem:prop41}
Assume additionally that the residue field $k$ is of characteristic $\neq 2$.
Then the canonical homomorphism $f\colon C(\rD_\ell, B, I) \to C(\rD_\ell, R, I)$ is surjective for $\ell \geq 7$. \end{lemma}
\begin{proof}
Writing the starting portion of the homology long exact sequence for the homomorphism $\St(\rD_\ell, R) \to \St(\rD_\ell, R/I)$ 
and using the isomorphisms of~\cref{lem:relativeH3} and~\cref{lem:lowerKgroups} we obtain the following commutative diagram:
\begin{equation*}\xymatrix{
 \KO_3(2\ell, B) \ar[r] \ar[d] & \KO_3(2\ell, k[X]) \ar[d]_{f'} \ar@{->>}[r] & \ar[d]_{f} C(\rD_\ell, B, I) \\
 \KO_3(2\ell, R) \ar[r]        & \KO_3(2\ell, k[X, X^{-1}]) \ar@{->>}[r]        & C(\rD_\ell, R, I).}\end{equation*}
By~\cref{Panin-stability} the homomorphism $f'$ can be identified with the canonical map $\GW_3^{[0]}(k[X]) \to \GW_3^{[0]}(k[X, X^{-1}])$.
By~\cref{bass-ft} $\GW_3^{[0]}(k[X, X^{-1}]) \cong \GW_3^{[0]}(k) \oplus \GW_2^{[3]}(k),$
but since the group $\GW_2^{[3]}(k)$ is trivial by~\cite[Lemma~2.2]{FRS12}, the homomorphism $f'$ (and hence $f$) is surjective.
\end{proof}

@@

\begin{externaltheorem} \label{thm41} Suppose that $2 \in A^\times$. Then for $\ell \geq 7$ the canonical homomorphism $\St(\rD_\ell, B) \to \St(\rD_\ell, R)$ is injective. \end{externaltheorem}
\begin{proof}
 Consider the following commutative diagram with exact rows obtained from~\eqref{LodayRelativization}:
\begin{equation*} \xymatrix{
 C(\rD_\ell, B, I) \ar[r]^{\lambda_B} \ar@{->>}[d]_{f} & \St(\rD_\ell, B, I) \ar[r]^{\mu_B} \ar[d]_{g} &
 \St(\rD_\ell, B) \ar[r]^{\nu_B} \ar[d]_{h} & \St(\rD_\ell, k[X]) \ar[d]_{i} \\
 C(\rD_\ell, R, I) \ar[r]^{\lambda_R}         & \St(\rD_\ell, R, I) \ar[r]^{\mu_R} \ar@{-->}[ur]^{t}&
 \St(\rD_\ell, R) \ar[r]^-{\nu_R}        & \St(\rD_\ell, k[X, X^{-1}]).
}\end{equation*}
The lifting $t$ in the above diagram is obtained from~\cref{lem:lemma32}.
Let $a$ be an element of $\Ker(h)$. Since $i$ is injective by~\cref{field-injectivity}, the element $a$
 also lies in $\Ker(\nu_B)$ and hence comes from some $b \in \St(\rD_\ell, B, I)$ via $\mu_B$.
Since $g(b) \in \Ker(\mu_R)$ there exists some $c \in C(\rD_\ell, R, I)$ such that $\lambda_R(c) = g(b)$. 
By~\cref{lem:prop41}, $f$ is surjective, therefore $c = f(d)$ for some $d \in C(\rD_\ell, R, I)$.
The required assertion now follows from the following calculation:
 \[ 1 = \mu_B\lambda_B(d) = tg\lambda_B(d) = t\lambda_Rf(d) =t(g(b)) = \mu_B(b) = a. \qedhere \]
\end{proof}

@@

For $i=1,\ldots,5$ denote by $\mathcal{G}_i$ the set of all generators of type K$i$. from~\cref{def:Kab}.
Notice that $\Phi \setminus \Psi$ contains the parabolic set of roots $(\Phi \setminus \Psi)\cap \Phi^+$.
Thus, by~\cref{thm:Stepanov} every element of $\overline{\St}(\Phi, A[X], XM[X])$ can be presented as a product of generators $\mathcal{G}_1$ and generators $\mathcal{G}' = \{ x_{\gamma}(Xf) \mid \gamma \in \Psi\}$. In turn, by~\cref{Kgen-strong} one can express $g$ as a product of these generators and generators $\mathcal{G}_5$.

@@

\subsection{Construction of a \texorpdfstring{$\St(\Phi, B)$}{St(B)}-torsor.} \label{sec:V-construction}
Throughout this subsection $\Phi$ denotes an arbitrary irreducible simply-laced root system of rank $\geq 3$, unless stated otherwise. 
\begin{df}\label{df:GM0_geq0}
Denote the subgroup $\overline{\St}(\Phi, A, M)$ by $G_M^0$. 
Consider the following commutative diagram, in which the homomorphisms $i_\pm$ and $j_\pm$ are induced by the natural ring homomorphisms $A \to A[X^\pm]$ and $A[X^\pm]\to R$:
\[\xymatrix{ G_M^0 \ar@{^{(}->}[rr]^{i_+^M} \ar@<-5pt>@{^{(}->}[dd]^{i_-^M} \ar@<-5pt>@{^{(}->}[rd] & & \overline{\St}(\Phi, A[X], M[X]) \ar[dd]_(.3){j_+^M} \ar@<-2.5pt>@{^{(}->}[rd] & \\ & \St(\Phi, A) \ar@{^{(}->}[rr]^(.3){i_+} \ar@<-2.5pt>@{^{(}->}[dd]^(.3){i_-} & & \St(\Phi, A[X]) \ar[dd]^{j_+} \\ \overline{\St}(\Phi, A[X\inv], M[X\inv]) \ar@<-2.5pt>@{^{(}->}[rd] \ar[rr]^(.35){j_-^M} & & \overline{\St}(\Phi, R, I) \ar@<-2.5pt>@{^{(}->}[rd] & \\ & \St(\Phi, A[X\inv]) \ar[rr]_{j_-} & & \St(\Phi, R)} \]
Denote by $\overline{G}^{\geq 0}_M$ the image of the homomorphism $j_+^M$. \end{df}
Notice that the homomorphism $j_+^Mi_+^M$ is split by the homomorphism $ev^*_{X=1}$ of evaluation at $1$, therefore $G_M^0$ can be considered as a subgroup of $\overline{G}^{\geq 0}_M$.

@@

Our next goal is to define for $a + Xf \in A + XM[X]$ the operator $T_\alpha(a + Xf) \colon \overline{V} \to \overline{V}$.
Let $(g, h, u)$ be an element of $V$. We define the value of $T_\alpha(a + Xf)$ on $(g, h, u)$ via the following identity:
\begin{equation} \label{T_leq0} T_\alpha(a + Xf) \cdot (g, h, u) = (x_\alpha(a + Xf) \cdot g, h \cdot x_{\alpha}(-a), u).  \end{equation}
Notice that $p(T_\alpha(a + Xf) \cdot (g, h, u)) = x_\alpha(Xf) \cdot {}^{x_\alpha(a)}p(g, h, u) \in \overline{G}_M^{\geq 0},$ therefore
 the right-hand side of~\eqref{T_leq0} lies in $V$. Notice that for $h_0 \in G^0_M$ one has
\[T_\alpha(a + Xf) \cdot ((g, h, u) \cdot h_0) = (T_\alpha(a + Xf) \cdot (g, h, u)) \cdot {}^{x_{\alpha}(a)}\!h_0,\]  
therefore~\eqref{T_leq0}, indeed, gives rise to a well-defined map $\overline{V} \to \overline{V}$.

@@

\begin{lemma} \label{R2_0_1} The operators $T_\alpha(aX^{-1})$ satisfy Steinberg relations $\mathrm{R2}_{0,1}$. \end{lemma}
\begin{proof} 
Let $a, b \in A$ and $\alpha, \beta$ be a pair of roots such that $\alpha + \beta \in \Phi$. Set $\epsilon = N_{\alpha, \beta},$ $c = 1 - \epsilon abm'$.

>>>>

\begin{lemma} \label{lem:prop41}
Assume additionally that the residue field $k$ is of characteristic $\neq 2$.
Then the canonical homomorphism $f\colon C(\rD_\ell, B, I) \to C(\rD_\ell, R, I)$ is surjective for $\ell \geq 7$. \end{lemma}
\begin{proof}
Writing the starting portion of the homology long exact sequence for the homomorphism $\St(\rD_\ell, R) \to \St(\rD_\ell, R/I)$ 
and using the isomorphisms of~\cref{lem:relativeH3} and~\cref{lem:lowerKgroups} we obtain the following commutative diagram:
\begin{equation*}\xymatrix{
 \KO_3(2\ell, B) \ar[r] \ar[d] & \KO_3(2\ell, k[X^{-1}]) \ar[d]_{f'} \ar@{->>}[r] & \ar[d]_{f} C(\rD_\ell, B, I) \\
 \KO_3(2\ell, R) \ar[r]        & \KO_3(2\ell, k[X, X^{-1}]) \ar@{->>}[r]        & C(\rD_\ell, R, I).}\end{equation*}
By~\cref{Panin-stability} the homomorphism $f'$ can be identified with the canonical map $\GW_3^{[0]}(k[X^{-1}]) \to \GW_3^{[0]}(k[X, X^{-1}])$.
By~\cref{bass-ft} $\GW_3^{[0]}(k[X, X^{-1}]) \cong \GW_3^{[0]}(k) \oplus \GW_2^{[3]}(k),$
but since the group $\GW_2^{[3]}(k)$ is trivial by~\cite[Lemma~2.2]{FRS12}, the homomorphism $f'$ (and hence $f$) is surjective.
\end{proof}

@@

\begin{externaltheorem} \label{thm41} Suppose that $2 \in A^\times$. Then for $\ell \geq 7$ the canonical homomorphism $\St(\rD_\ell, B) \to \St(\rD_\ell, R)$ is injective. \end{externaltheorem}
\begin{proof}
 Consider the following commutative diagram with exact rows obtained from~\eqref{LodayRelativization}:
\begin{equation*} \xymatrix{
 C(\rD_\ell, B, I) \ar[r]^{\lambda_B} \ar@{->>}[d]_{f} & \St(\rD_\ell, B, I) \ar[r]^{\mu_B} \ar[d]_{g} &
 \St(\rD_\ell, B) \ar[r]^{\nu_B} \ar[d]_{h} & \St(\rD_\ell, k[X^{-1}]) \ar[d]_{i} \\
 C(\rD_\ell, R, I) \ar[r]^{\lambda_R}         & \St(\rD_\ell, R, I) \ar[r]^{\mu_R} \ar@{-->}[ur]^{t}&
 \St(\rD_\ell, R) \ar[r]^-{\nu_R}        & \St(\rD_\ell, k[X, X^{-1}]).
}\end{equation*}
The lifting $t$ in the above diagram is obtained from~\cref{lem:lemma32}.
Let $a$ be an element of $\Ker(h)$. Since $i$ is injective by~\cref{field-injectivity}, the element $a$
 also lies in $\Ker(\nu_B)$ and hence comes from some $b \in \St(\rD_\ell, B, I)$ via $\mu_B$.
Since $g(b) \in \Ker(\mu_R)$ there exists some $c \in C(\rD_\ell, R, I)$ such that $\lambda_R(c) = g(b)$. 
By~\cref{lem:prop41}, $f$ is surjective, therefore $c = f(d)$ for some $d \in C(\rD_\ell, R, I)$.
The required assertion now follows from the following calculation:
 \[ 1 = \mu_B\lambda_B(d) = tg\lambda_B(d) = t\lambda_Rf(d) =t(g(b)) = \mu_B(b) = a. \qedhere \]
\end{proof}

@@

For $i=1,\ldots,5$ denote by $\mathcal{G}_i$ the set of all generators of type K$i$ from~\cref{def:Kab}.
Notice that $\Phi \setminus \Psi$ contains the parabolic set of roots $(\Phi \setminus \Psi)\cap \Phi^+$.
Thus, by~\cref{thm:Stepanov} every element of $\overline{\St}(\Phi, A[X], XM[X])$ can be presented as a product of generators $\mathcal{G}_1$ and generators $\mathcal{G}' = \{ x_{\gamma}(Xf) \mid \gamma \in \Psi\}$. In turn, by~\cref{Kgen-strong} one can express $g$ as a product of these generators and generators $\mathcal{G}_5$.

@@

\subsection{Construction of a \texorpdfstring{$\St(\Phi, B)$}{St(B)}-torsor.} \label{sec:V-construction}
Throughout this subsection $\Phi$ denotes an arbitrary irreducible simply-laced root system of rank $\geq 3$, unless stated otherwise. 
\begin{df}\label{df:GM0_geq0}
Denote the subgroup $\overline{\St}(\Phi, A, M)$ by $G_M^0$. 
Consider the following commutative diagram, in which the homomorphisms $i_\pm$ and $j_\pm$ are induced by the natural ring homomorphisms $A \to A[X^\pm]$ and $A[X^\pm]\to R$:
\[\xymatrix{ G_M^0 \ar@{^{(}->}[rr]^{i_+^M} \ar@<-5pt>@{^{(}->}[dd]^{i_-^M} \ar@<-5pt>@{^{(}->}[rd] & & \overline{\St}(\Phi, A[X], M[X]) \ar[dd]_(.3){j_+^M} \ar@<-2.5pt>@{^{(}->}[rd] & \\ & \St(\Phi, A) \ar@{^{(}->}[rr]^(.3){i_+} \ar@<-2.5pt>@{^{(}->}[dd]^(.3){i_-} & & \St(\Phi, A[X]) \ar[dd]^{j_+} \\ \overline{\St}(\Phi, A[X\inv], M[X\inv]) \ar@<-2.5pt>@{^{(}->}[rd] \ar[rr]^(.35){j_-^M} & & \overline{\St}(\Phi, R, I) \ar@<-2.5pt>@{^{(}->}[rd] & \\ & \St(\Phi, A[X\inv]) \ar[rr]_{j_-} & & \St(\Phi, R).} \]
Denote by $\overline{G}^{\geq 0}_M$ the image of the homomorphism $j_+^M$. \end{df}
Notice that the homomorphism $j_+^Mi_+^M$ is split by the homomorphism $ev^*_{X=1}$ of evaluation at $1$, therefore $G_M^0$ can be considered as a subgroup of $\overline{G}^{\geq 0}_M$.

@@

Our next goal is to define for $a + Xf \in A + XM[X]$ the operator $T_\alpha(a + Xf) \colon \overline{V} \to \overline{V}$.
Let $(g, h, u)$ be an element of $V$. We define the value of $T_\alpha(a + Xf)$ on $(g, h, u)$ via the following identity:
\begin{equation} \label{T_leq0} T_\alpha(a + Xf) \cdot (g, h, u) = (x_\alpha(a + Xf) \cdot g, h \cdot x_{\alpha}(-a), u).  \end{equation}
Notice that $p(T_\alpha(a + Xf) \cdot (g, h, u)) = x_\alpha(Xf) \cdot {}^{x_\alpha(a)}p(g, h, u) \in \overline{G}_M^{\geq 0},$ therefore
 the right-hand side of~\eqref{T_leq0} lies in $V$. Notice that for $h_0 \in G^0_M$ one has
\[T_\alpha(a + Xf) \cdot ((g, h, u) \cdot h_0) = (T_\alpha(a + Xf) \cdot (g, h, u)) \cdot {}^{x_{\alpha}(a)}\!h_0,\]  
therefore~\eqref{T_leq0} indeed gives rise to a well-defined map $\overline{V} \to \overline{V}$.

@@

\begin{lemma} \label{R2_0_1} The operators $T_\alpha$ satisfy Steinberg relations $\mathrm{R2}_{0,1}$. \end{lemma}
\begin{proof} 
Let $a, b \in A$ and $\alpha, \beta$ be a pair of roots such that $\alpha + \beta \in \Phi$. Set $\epsilon = N_{\alpha, \beta},$ $c = 1 - \epsilon abm'$.
